---
title: "Boulder Festival Insights Pre-processing and EDA"
output: html_document
---

Every year, the city of Boulder, CO hosts a major festival. The festival is meant to attract tourists from all around the country bolstering the marketing of Boulder while serving as an economic stimulant. This document takes a look at a survey conducted during the actual festival and aims to report what spending levels looked like at the festival as just one gauge of what economic benefits the festival brings to the city.

__Data Source:__

* assign_2.csv

__Below is a list of variables that were of importance:__

* X - Survey identifier; unique individual attending the festival
* visits - Number of times someone has visited Festival in the past
* days_attend - Number of days attended
* hours_attend - Number of hours attended per day, on average
* gender - The reported gender of the survey participant
* age - The reported age of the participant
* zip - The reported home zip code of the participant
* extend - Whether the participant extended their trip due to Festival


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

#### Reading in festival data 

```{r}
festival_data <- read_csv('assign_2.csv')
```

#### Creating a separate data set of all of the duplicate observations

```{r}
festival_data_1 <- festival_data %>%
  filter(X == 1)
```

#### Finding the number of duplicates 

```{r}
n_repeated_1 <- festival_data_1 %>%
  nrow()
```
```{r, echo=FALSE}
n_repeated_1
```

#### Creating a separate data set that contains no duplicate values

```{r}
festival_data_unique <- festival_data %>%
  distinct()
```

#### Creating a new column that takes on a binary value 1 if an individual extended their trip and 0 otherwise

```{r}
festival_data_unique <- festival_data_unique %>%
  mutate(extend_b = ifelse(extend == 'Yes', 1, 0))
```

#### Finding if those who extended their stay also spent more time each day at the festival and how many more (or less) hours they attended

```{r}
extend_1_hours <- festival_data_unique %>%
  filter(extend_b==1) %>%
  summarise(mean(hours_attend, na.rm = T))

extend_0_hours <- festival_data_unique %>%
  filter(extend_b==0) %>%
  summarise(mean(hours_attend, na.rm = T))

```

Amount of hours individuals who extended their stay stayed for:

```{r}
extend_more_hours <- extend_1_hours - extend_0_hours
```
```{r, echo=FALSE}
show(extend_more_hours)
```

#### Finding the distinct intervals of spending by people at the festival

```{r}
f_data_spend_entertainment_total_amounts <- festival_data_unique %>%
  distinct(spend_entertainment_total) %>%
  drop_na()
```
```{r, echo=FALSE}
show(f_data_spend_entertainment_total_amounts)
```

#### Creating a table that counts the amount of people that spent over $100

```{r}
spend_more_100 <- c("$500-$999", "$200-$499")

festival_data_unique <- festival_data_unique %>% 
  mutate(s_entertainment_total_more_100 = ifelse(spend_entertainment_total %in% spend_more_100, 1, 0))
```
```{r, echo=FALSE}
show(festival_data_unique)
```

#### Calculating the proportion of women who spent more than $100 on entertainment

```{r}
spend_entertainment_total_more_100_women <- festival_data_unique %>%
  filter(gender=='Female') %>%
  summarise(spend_entertainment_total_more_100_women = mean(s_entertainment_total_more_100))
```
```{r, echo=FALSE}
show(spend_entertainment_total_more_100_women)
```

#### Creating a tibble with all possible cominations of gender and age and the corresponding proportions of who spent more than $100 on entertaining

```{r}
spend_entertainment_total_more_100_demo <- festival_data_unique %>%
  group_by(gender, age) %>%
  summarise(se_t_m_100_demo = mean(s_entertainment_total_more_100, na.rm = T))
```
```{r, echo=FALSE}
show(spend_entertainment_total_more_100_demo)
```

#### Creating a table of all the age ranges and the proportion of individuals that went to the festival that fall in those intervals

```{r}
ages <- festival_data_unique %>%
  group_by(age) %>%
  count() 

ages <- ages %>%
  mutate(proportion = n/287)
```
```{r, echo=FALSE}
show(ages)
```

#### Counting the amount of times people have visited the festival

```{r}
visits <- festival_data_unique %>%
  group_by(visits) %>%
  count(sort = T)
```
```{r, echo=FALSE}
show(visits)
```
 
#### Finding the amount of people that extended their stay and also spent more than $100 dollars on entertainment at the festival 

```{r}
extend_hours_over_100 <- festival_data_unique %>%
  filter(extend_b==1, spend_entertainment_total %in% spend_more_100) %>%
  nrow()
```
```{r, echo=FALSE}
show(extend_hours_over_100)
```

#### Finding the amount of people that didn't extend their stay but still spent more than $100 dollars on entertainment

```{r}
nonextend_hours_over_100 <- festival_data_unique %>%
  filter(extend_b==0, spend_entertainment_total %in% spend_more_100) %>%
  nrow()

```
```{r, echo=FALSE}
show(nonextend_hours_over_100)
```

#### Finding the amount of people that extended their stay at the festival

```{r}
extended_hours <- festival_data_unique %>%
  filter(extend_b==1) %>%
  nrow()
```
```{r, echo=FALSE}
show(extended_hours)
```

#### Finding the amount of people that didn't extend their stay at the festival

```{r}
nonextended_hours <- festival_data_unique %>%
  filter(extend_b==0) %>%
  nrow()
```
```{r, echo=FALSE}
show(nonextended_hours)
```

#### Finding the proportion of people that spent over $100 when they extended their stay and the proportion of people that spent over $100 even when they didn't extended their stay

```{r}
prop_extend_100 <- extend_hours_over_100 / extended_hours
prop_nonextend_100 <- nonextend_hours_over_100 / nonextended_hours
```
```{r, echo=FALSE}
show(prop_extend_100)
show(prop_nonextend_100)
```

